---
title: "introduction"
author: "Seonjun Park"
data: "2025-05-05"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE, 
  collapse = TRUE,
  comment = "#>"
)
```

### Setup.

```{r setup}
library(enCORE)
library(dplyr)
library(dplyr)
library(ArchR)
library(parallel)
library(motifmatchr)
library(chromVARmotifs)
library(BSgenome.Hsapiens.UCSC.hg38)
setwd("~/PSJ/enCORE_dev")
h5disableFileLocking()
addArchRThreads(13)
addArchRGenome("hg38")
data("human_pwms_v2")
```

### Obtain co-accessibility information from ArchR object.

```{r}
proj4 <- readRDS("~/PSJ/test_enCORE/Save-Proj_r4/Save-ArchR-Project.rds")

proj5 <- addCoAccessibility(
  ArchRProj = proj4,
  reducedDims = "IterativeLSI",
  maxDist = 500000
)

cA <- getCoAccessibility(
  ArchRProj = proj5,
  corCutOff = 0.2,
  resolution = 1,
  returnLoops = FALSE
)
```

### Add cluster information for performing enCORE.
enCORE can only recognize the column metadata information called Clusters2.

```{r}
proj5$Clusters2 <- mapLabels(proj5$Sample, newLabels = remapClust, oldLabels = names(remapClust))
```

### Extract initial enhancer candidates.
After running the following code, two required files were generated in output_dir.

```{r}
data_lump_enCORE <- Extract_initial_enhancer_candidates(proj_atac = proj5, data_coacc = cA, output_dir = "~/PSJ/test_enCORE", organism = "hg38")
```

### Calculate gABC scores.
To perform the following code properly, .gtf file must be located in the previous output_dir (== "~/PSJ/test_enCORE"). After running the following code, gABC scores per cluster were generated in output_dir (== "~/PSJ/test_enCORE/results_gABC_cr_f").

```{r}
data_lump_enCORE <- Calculate_gABC_score(data_lump_enCORE = data_lump_enCORE, n_col = 5, STARE_dir = "~/anaconda3/bin", output_dir = "~/PSJ/test_enCORE/results_gABC_cr_f")
```

### Add TF motif information to ArchR object.
Add CISBP motif and chromVAR deviation information to ArchR object.

```{r}
proj6 <- addMotifAnnotations(ArchRProj = proj5, motifPWMs = human_pwms_v2, name = "Motif")
proj6 <- addBgdPeaks(proj6)
proj6 <- addDeviationsMatrix(
  ArchRProj = proj6,
  peakAnnotation = "Motif",
  force = TRUE,
  threads = 10
)
```

### Determine TF weight threshold.
Default threshold is 0.675. If not, automatically determine TF weight threshold. If you want to use your own theshold, you should set use_default_thres = TRUE & predefined_thres = [your threshold].

```{r}
data_lump_enCORE <- Determine_TF_weight_threshold(proj_atac = proj6, data_lump_enCORE = data_lump_enCORE, list_cluster = c("CRC", "Normal"), use_default_thres = FALSE)
```

### Extract CORE.
Write CORE regions into output_dir. If you use option = "potential", only raw CORE. However, if you use option = "active", you can obtain both regions, CORE from the potential (raw) option and the active (filtered) option.

```{r}
data_lump_enCORE <- Distill_CORE_per_cluster(proj_atac = proj6, data_lump_enCORE = data_lump_enCORE, output_dir = "~/PSJ/test_enCORE/test_filtered", option = "active")
```
